Supporting Information

### 2017-04-06 code/rmarkdown/figures.Rmd
massive overhaul - packrat to manage r packages - reformat code according to lintr - fix book references

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
```

```{r, include=FALSE}
## initialization
# load .rda file
checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)

## prelimary processing
# redefine file paths for rasters
for (i in seq_len(nlayers(cs1.bioclim.RAST)))
	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'raptr-manuscript')[[1]][[2]])

# remove small islands from australia basemap
aus.sub.PLY <- disaggregate(aus.PLY)
aus.sub.PLY <- SpatialPolygonsDataFrame(aus.sub.PLY, data=data.frame(size=gArea(aus.sub.PLY, byid=TRUE)))
aus.sub.PLY <- aus.sub.PLY[aus.sub.PLY$size >= quantile(aus.sub.PLY$size, 0.9, names=FALSE),]

# fortify polygons
sim.pus.FPLY <- fortify(sim.pus)
aus.FPLY <- fortify(aus.sub.PLY)
cs1.pus.FPLY <- fortify(cs1.pus)
cs2.grid.FPLY <- fortify(cs2.grid.PLY)
cs2.grid.sub.FPLY <- fortify(cs2.grid.sub.PLY)

# store countries
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
	)
,] %>% spFortify

column.width <- 2.75591

@@

```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap="Attribute space example. This environmental attribute space has dimensions relating to annual temperature ($^{\\circ}$C) and rainfall (mm). Letters denote the environmental conditions associated with the geographic locations where four hypothetical populations are found. Points represent demand points. In this space, populations close to each other inhabit similar environmental conditions."}
## generate data
# population data
populations <- data.frame(
	population=LETTERS[1:4],
	temperature=c(35,34,31,21.6),
	rainfall=c(4.5,5.3,9.5,5.5)   
)

# demand point data
make.dp <- function(x, n, pad.percent=0.2) {
	padding<-diff(range(x))*pad.percent
	return(seq(min(x)-padding, max(x)+padding, length.out=n))
}
demand.points <- expand.grid(
	temperature=make.dp(populations$temperature, n=4, pad.percent=0.2),
	rainfall=make.dp(populations$rainfall, n=4, pad.percent=0.2)
)

## plot data
ggplot(aes(x=rainfall, y=temperature), data=populations) +
	geom_point(data=demand.points, fill='gray70') +
	geom_text(aes(label=population), size=6) +
	theme_classic() +
	ylab(expression('Temperature ('*degree*'C)')) +
	xlab('Rainfall (mm)') +
	theme(axis.line.x=element_line(), axis.line.y=element_line())
```

```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.6, fig.width=column.width, fig.cap='Distributions of three simulated species. Squares denote planning units. Colors indicate probability of occupancy.'}
# prepare data
sim.pus.occ.FPLY <- ldply(
	seq_len(nrow(sim.species)),
	function(i) {
		curr.DF <- sim.pus.FPLY
		curr.DF$species <- sim.species$name[i]
		curr.DF$Probability <- sim.dps[[i]]@weights[as.numeric(curr.DF$id)]
		return(curr.DF)
	}
)
sim.pus.occ.FPLY$species <- gsub('\n', ' ', as.character(sim.pus.occ.FPLY$species))
sim.pus.occ.FPLY$species <- factor(as.character(sim.pus.occ.FPLY$species), levels=gsub('\n', ' ', sim.species$name)) 
# make letter data.frame
sim.pus.occ.letters.DF <- data.frame(
	x=min(sim.pus.occ.FPLY$long) + 1.5,
	y=max(sim.pus.occ.FPLY$lat) - 1.5,
	species=unique(sim.pus.occ.FPLY$species),
	letter=paste0('(', letters[seq_along(sim.species$name)], ')')
)
# make plot
ggplot() +
	geom_polygon(data=sim.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Probability), color='black') +
	theme_classic() +
	geom_label(data=sim.pus.occ.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=2) +
	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', size=7),
    axis.title = element_blank(),
    legend.title = element_text(color='black', size=7),
		legend.text = element_text(color='black', size=7),
		legend.margin = unit(0, 'cm'),
		legend.position='bottom') +
	facet_wrap(~ species, ncol=3)
```

```{r, echo=FALSE, fig.height=3.5, fig.width=2*column.width, fig.cap=paste0('Prioritizations for the simulation study. Each panel shows a prioritization generated for a single species using a set of parameters. Squares denote planning units. Planning units shown in dark green units were selected for protection. Each row of panels show prioritizations generated for a different species. Each column of panels corresponds to a different set of parameters used to generate the prioritization.')}
# prepare data
sim.pus.sel.FPLY <- ldply(
	seq_len(nrow(sim.spp.DF)),
	function(i) {
		curr.DF <- sim.pus.FPLY %>% mutate(species=sim.spp.DF$species[i], prioritisation=sim.spp.DF$prioritisation[i])
		curr.selection <- selections(sim.prioritisations[[as.character(sim.spp.DF$species[i])]][[as.character(sim.spp.DF$prioritisation[i])]])[as.numeric(curr.DF$id)]
		curr.DF$Selection <-  c('discarded','selected')[curr.selection+1]
		return(curr.DF)
	}
)
sim.pus.sel.FPLY$prioritisation <- factor(as.character(sim.pus.sel.FPLY$prioritisation),
	levels=c('Amount\ntargets', 'Amount &\nspace targets', 'Amount target\nand BLM', 'Amount & space\ntargets and BLM'))
# prepare plotting data
sim.letter.DF2 <- sim.spp.DF %>%
  mutate(letter=paste0('(', letters[seq_along(species)],')'),
         long=min(sim.pus.sel.FPLY$long) + 1.6,
         lat=max(sim.pus.sel.FPLY$lat) - 1.5)
# make plot
ggplot() +
	geom_polygon(data=sim.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Selection), color='black') +
	geom_label(data=sim.letter.DF2, aes(label=letter, x=long, y=lat), color='black', fill=alpha('white', 0.8), size=5) +
	theme_classic() +
	theme(
		axis.ticks=element_blank(), axis.ticks.x=element_blank(), axis.ticks.y=element_blank(),
		axis.text=element_blank(), axis.text.x=element_blank(), axis.text.y=element_blank(),
		axis.line=element_blank(), axis.line.x=element_blank(), axis.line.y=element_blank(),
		plot.background=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		panel.background=element_blank(), panel.border=element_blank(), legend.position='none',
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white'),
		axis.title=element_blank()) +
	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
	facet_grid(species ~ prioritisation)
```

```{r, echo=FALSE, fig.height=1.9, fig.width=column.width, fig.cap='Distribution of climatic variation across Queensland, Australia. Squares denote planning units. Panels show the two main gradients of climatic variation.'}
# prepare data
cs1.pca.FPTS <- ldply(
	seq_len(cs1.params.LST[[MODE]]$n.pc.layers),
	function(i) {
		curr.DF <- cs1.bioclim.RAST[[i]] %>% 
			aggregate(10) %>% 
			rasterToPoints() %>% 
			as.data.frame()
		names(curr.DF)[3] <- 'Score'
		curr.DF$pc <- paste0('PC ',i)
		return(curr.DF)
	}
)
# make plots
p1 <- ggplot() +
	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
	geom_raster(data=filter(cs1.pca.FPTS, pc==unique(pc)[1]), aes(x=x, y=y, fill=Score)) +
	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black', size=0.05) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white'),
		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.height=unit(1.5, 'mm'), legend.key.width=unit(2, 'mm')
	) +
  coord_cartesian(xlim=buffered.range(cs1.pus@bbox[1,], 0.025), ylim=buffered.range(cs1.pus@bbox[2,], 0.025), expand=FALSE) +
  scale_fill_continuous(breaks=pretty(filter(cs1.pca.FPTS, pc==unique(pc)[1])$Score, n=3, min.n=2)) +
	facet_wrap(~ pc)
p2 <- ggplot() +
	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
	geom_raster(data=filter(cs1.pca.FPTS, pc==unique(pc)[2]), aes(x=x, y=y, fill=Score)) +
	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black', size=0.05) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white'),
		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.height=unit(1.5, 'mm'), legend.key.width=unit(2, 'mm')
		) +
  coord_cartesian(xlim=buffered.range(cs1.pus@bbox[1,], 0.025), ylim=buffered.range(cs1.pus@bbox[2,], 0.025), expand=FALSE) +
	scale_fill_continuous(breaks=pretty(filter(cs1.pca.FPTS, pc==unique(pc)[2])$Score, n=2, min.n=2)) +
	facet_wrap(~ pc)
# render plots
grid.arrange(p1, p2, ncol=2)
```

```{r, echo=FALSE, fig.height=3.8, fig.width=column.width, fig.cap='Distribution of the species used in the first case study. Each panel shows the geographic range for a different species. Squares denote planning units.'}
# prepare data
cs1.pus.occ.FPLY <- ldply(
	seq_along(cs1.params.LST[[MODE]]$common.names),
	function(i) {
		curr.DF <- cs1.pus.FPLY
		curr.DF$species <- cs1.params.LST[[MODE]]$common.names[i]
		curr.DF$Occupancy <- 'absent'
		curr.DF$Occupancy[which(curr.DF$id %in% filter(cs1.ru@data@pu.species.probabilities, species==i)$pu)] <- 'present'
		return(curr.DF)
	}
)

cs1.pus.occ.FPLY$species <- factor(as.character(cs1.pus.occ.FPLY$species), levels=cs1.params.LST[[MODE]]$common.names)
# make letter data.frame
cs1.pus.occ.letters.DF <- data.frame(
	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.08)),
	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.95)),
	species=unique(cs1.pus.occ.FPLY$species),
	letter=paste0('(', letters[seq_along(unique(cs1.pus.occ.FPLY$species))], ')')
)
# make plots
ggplot() +
	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70', size=0.05) +
	geom_polygon(data=cs1.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), color='black', size=0.05) +
	geom_label(data=cs1.pus.occ.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', size=7),
		axis.title=element_blank(), legend.position='none'
	) +
	scale_fill_manual(name='Occupancy', values=c('absent'='#132B43', 'present'='#56B1F7')) +
  coord_cartesian(xlim=buffered.range(cs1.pus@bbox[1,], 0.025), ylim=buffered.range(cs1.pus@bbox[2,], 0.025), expand=FALSE) +
	facet_wrap(~ species, ncol=2)
```

```{r, echo=FALSE, fig.height=1.9, fig.width=column.width, fig.cap=paste0('Prioritizations for the first case study. See Fig. 3 caption for conventions. Panel (a) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets. Panel (b) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs1.params.LST[[MODE]]$space.target*100,' % niche targets.')}
# prepare data
cs1.pus.sel.FPLY <- ldply(
	names(cs1.prioritisations),
	function(x) {
		curr.DF <- cs1.pus.FPLY
		curr.DF$Prioritisation <- x
		curr.DF$Solution <- 'discarded'
		curr.DF$Solution[curr.DF$id %in% which(as.logical(selections(cs1.prioritisations[[x]])))] <- 'selected'
		return(curr.DF)
	}
)
cs1.pus.sel.FPLY$Prioritisation <- factor(as.character(cs1.pus.sel.FPLY$Prioritisation), levels=names(cs1.prioritisations))
# make letter data.frame
cs1.pus.sel.letters.DF <- data.frame(
	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.08)),
	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.95)),
	Prioritisation=names(cs1.prioritisations),
	letter=paste0('(', letters[1:2], ')')
)
# make plots
ggplot() +
	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70', size=0.05) +
	geom_polygon(data=cs1.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Solution), color='black', size=0.05) +
	geom_label(data=cs1.pus.sel.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', size=8),
		axis.title=element_blank(), legend.position='none'
	) +
	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
	coord_cartesian(xlim=buffered.range(cs1.pus@bbox[1,], 0.025), ylim=buffered.range(cs1.pus@bbox[2,], 0.025), expand=FALSE) +
	facet_wrap(~ Prioritisation, ncol=2)
```

```{r, echo=FALSE, fig.height=2.2, fig.width=column.width, fig.cap=paste0('Data used for the second case study. Squares denote planning units. Panel (a) shows all grid cells surveyed by the IntraBioDiv project. Grid cells occupied by ',tolower(cs2.params.LST[[MODE]]$common.name),' are shown in bright blue. The subsequent panels contain only show occupied grid cells. Panel (b) shows the opportunity cost of each planning unit (estimated as the total human population density). Panels (c--d) show the spatial distribution of the ordinations describing genetic variation. These values describe the typical genetic characteristics of individuals in each planning unit. Planning units with similar values/colors contain individuals with similar loci polymorphisms.')}
# prepare full grid data
cs2.spp.grid.FPLY <- cs2.grid.FPLY
cs2.spp.grid.FPLY$Title <- 'Occupancy'
cs2.spp.grid.FPLY$Occupancy <- 'absent'
cs2.spp.grid.FPLY$Occupancy[which(cs2.grid.PLY@data[[cs2.params.LST[[MODE]]$species.name]][as.numeric(cs2.spp.grid.FPLY$id)]==1)] <- 'present'

cs2.spp.grid.letters.DF <- data.frame(
	x=unname(cs2.grid.PLY@bbox[1,1] + (diff(cs2.grid.PLY@bbox[1,])*0.07)),
	y=unname(cs2.grid.PLY@bbox[2,1] + (diff(cs2.grid.PLY@bbox[2,])*0.875)),
	letter=('(a)')
)

>>>>

```{r, include = FALSE}
# set wd
knitr::opts_knit$set(root.dir = normalizePath("../.."))
```

```{r, include = FALSE}
## initialization
# load .rda file
try(session::restore.session("data/final/results.rda"))

## prelimary processing
# redefine file paths for rasters
for (i in seq_len(raster::nlayers(cs1_bioclim_raster)))
  cs1_bioclim_raster@layers[[i]]@file@name  <- paste0(getwd(),
    strsplit(cs1_bioclim_raster@layers[[i]]@file@name,
             "raptr-manuscript")[[1]][[2]])

# remove small islands from australia basemap
aus_sub_polygons <- sp::disaggregate(aus_polygons)
aus_sub_polygons <- aus_sub_polygons %>%
                    sp::SpatialPolygonsDataFrame(
                      data = data.frame(size = rgeos::gArea(aus_sub_polygons,
                                                            byid = TRUE)))
aus_sub_polygons <- aus_sub_polygons[aus_sub_polygons$size >=
  quantile(aus_sub_polygons$size, 0.9, names=FALSE),]

# fortify polygons
sim_pus_fpolygons <- ggplot2::fortify(sim_pus)
aus_fpolygons <- ggplot2::fortify(aus_sub_polygons)
cs1_pus_fpolygons <- ggplot2::fortify(cs1_pus)
cs2_grid_fpolygons <- ggplot2::fortify(cs2_grid_polygons)
cs2_grid_sub_fpolygons <- ggplot2::fortify(cs2_grid_sub_polygons)

# store countries
data(countriesHigh, package = "rworldxtra")
countries_fpolygons <- countriesHigh[
  countriesHigh$ADMIN %in% c(
    "Italy", "Switzerland", "France", "Austria",
    "Germany", "Slovenia", "Croatia", "Hungary",
    "Monaco", "Germany", "Slovakia", "Czech Republic"
  ),] %>% sp_fortify

column_width <- 2.75591

@@

```{r, echo = FALSE, message = FALSE, fig.height = 2, fig.width = 2, fig.cap = "Attribute space example. This environmental attribute space has dimensions relating to annual temperature ($^{\\circ}$C) and rainfall (mm). Letters denote the environmental conditions associated with the geographic locations where four hypothetical populations are found. Points represent demand points. In this space, populations close to each other inhabit similar environmental conditions."}
## generate data
# population data
populations <- data.frame(population = LETTERS[1:4],
                          temperature = c(35, 34, 31, 21.6),
                          rainfall = c(4.5, 5.3, 9.5, 5.5))

# demand point data
make_dp <- function(x, n, pad.percent = 0.2) {
  padding <- diff(range(x)) * pad.percent
  return(seq(min(x) - padding, max(x) + padding, length.out = n))
}
demand.points <- expand.grid(
  temperature = make_dp(populations$temperature, n = 4, pad.percent = 0.2),
  rainfall = make_dp(populations$rainfall, n = 4, pad.percent = 0.2))

## plot data
ggplot2::ggplot(ggplot2::aes(x = rainfall, y = temperature),
                data = populations) +
  ggplot2::geom_point(data = demand.points, fill = "gray70") +
  ggplot2::geom_text(ggplot2::aes(label = population), size = 6) +
  ggplot2::theme_classic() +
  ggplot2::ylab(expression("Temperature ("*degree*"C)")) +
  ggplot2::xlab("Rainfall (mm)") +
  ggplot2::theme(axis.line.x = ggplot2::element_line(),
                 axis.line.y = ggplot2::element_line())
```

```{r, echo = FALSE, messages = FALSE, results = "hide", fig.height = 1.6, fig.width = column_width, fig.cap = "Distributions of three simulated species. Squares denote planning units. Colors indicate probability of occupancy."}
# prepare data
sim_pus_occ_fpolygons <- plyr::ldply(
  seq_len(nrow(sim_species)),
  function(i) {
    curr_data <- sim_pus_fpolygons
    curr_data$species <- sim_species$name[i]
    curr_data$Probability <- sim_demand_points[[i]]@weights[
      as.numeric(curr_data$id)]
    return(curr_data)
  }
)
sim_pus_occ_fpolygons$species <- gsub("\n", " ",
  as.character(sim_pus_occ_fpolygons$species))
sim_pus_occ_fpolygons$species <- factor(as.character(
  sim_pus_occ_fpolygons$species), levels = gsub("\n", " ", sim_species$name))
# make letter data.frame
sim_pus_occ_letters <- data.frame(
  x = min(sim_pus_occ_fpolygons$long) + 1.5,
  y = max(sim_pus_occ_fpolygons$lat) - 1.5,
  species = unique(sim_pus_occ_fpolygons$species),
  letter=paste0("(", letters[seq_along(sim_species$name)], ")"))
# make plot
ggplot2::ggplot() +
  ggplot2::geom_polygon(data = sim_pus_occ_fpolygons,
                        ggplot2::aes(x = long, y = lat,  group = group,
                                     fill = Probability),
                        color="black") +
  ggplot2::theme_classic() +
  ggplot2::geom_label(data = sim_pus_occ_letters,
                     ggplot2::aes(label = letter, x = x, y =y ), color="black",
                     fill = scales::alpha("white", 0.8), size = 2) +
  ggplot2::theme(axis.ticks = ggplot2::element_blank(),
                axis.text = ggplot2::element_blank(),
                axis.line = ggplot2::element_blank(),
                strip.background = ggplot2::element_rect(fill = "grey20"),
                strip.text = ggplot2::element_text(color = "white", size = 7),
                axis.title = ggplot2::element_blank(),
                legend.text = ggplot2::element_text(color = "black", size = 7),
                legend.margin = ggplot2::unit(0, "cm"),
                legend.position = "bottom") +
  ggplot2::facet_wrap(~ species, ncol = 3)
```

```{r, echo = FALSE, fig.height = 3.5, fig.width = 2 * column_width, fig.cap = paste0("Prioritizations for the simulation study. Each panel shows a prioritization generated for a single species using a set of parameters. Squares denote planning units. Dark green planning units were selected for protection. Each row of panels show prioritizations generated for a different species. Each column of panels corresponds to a different set of parameters used to generate the prioritization.")}
# prepare data
sim_pus_sel_fpolygons <- plyr::ldply(seq_len(nrow(sim_spp_results)), function(i) {
    curr_data <- sim_pus_fpolygons %>%
                 dplyr::mutate(species = sim_spp_results$species[i],
                               prioritisation =
                                sim_spp_results$prioritisation[i])
    curr_selection <- sim_prioritisations %>%
                      `[[`(as.character(sim_spp_results$species[i])) %>%
                      `[[`(as.character(sim_spp_results$prioritisation[i])) %>%
                      raptr::selections() %>%
                      `[`(as.numeric(curr_data$id))
    curr_data$Selection <-  c("discarded","selected")[curr_selection+1]
    return(curr_data)
})
sim_pus_sel_fpolygons$prioritisation <- factor(
  as.character(sim_pus_sel_fpolygons$prioritisation),
  levels = c("Amount\ntargets", "Amount &\nspace targets",
             "Amount target\nand BLM", "Amount & space\ntargets and BLM"))
# prepare plotting data
sim_letter_data2 <- sim_spp_results %>%
  dplyr::mutate(letter = paste0("(", letters[seq_along(species)],")"),
                long = min(sim_pus_sel_fpolygons$long) + 1.6,
                lat = max(sim_pus_sel_fpolygons$lat) - 1.5)
# make plot
ggplot2::ggplot() +
  ggplot2::geom_polygon(data = sim_pus_sel_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group,
                                     fill = Selection),
                         color="black") +
  ggplot2::geom_label(data = sim_letter_data2,
                      ggplot2::aes(label = letter, x = long, y = lat),
                      color = "black", fill = scales::alpha("white", 0.8),
                      size=5) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.ticks.x = ggplot2::element_blank(),
    axis.ticks.y = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_blank(),
    axis.text.y = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    axis.line.x = ggplot2::element_blank(),
    axis.line.y = ggplot2::element_blank(),
    plot.background = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    panel.background = ggplot2::element_blank(),
    panel.border = ggplot2::element_blank(),
    legend.position="none",
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white"),
    axis.title = ggplot2::element_blank()) +
  ggplot2::scale_fill_manual(name="Solution",
                             values = c("discarded"="#f7fcfd",
                                        "selected"="#00441b")) +
  ggplot2::facet_grid(species ~ prioritisation)
```

```{r, echo = FALSE, fig.height = 1.9, fig.width = column_width, fig.cap = "Two main gradients of climatic variation across Queensland, Australia. Polygons denote planning units."}
# prepare data
cs1_pca_fpoints <- plyr::ldply(seq_len(cs1_parameters$number_pc_layers),
  function(i) {
    curr_data <- cs1_bioclim_raster[[i]] %>%
                     raster::aggregate(10) %>%
                     raster::rasterToPoints() %>%
                     as.data.frame()
    names(curr_data)[3] <- "Score"
    curr_data$pc <- paste0("PC ",i)
    return(curr_data)
})
# make plots
p1 <- ggplot2::ggplot() +
  ggplot2::geom_polygon(data = aus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = "grey85", color = "grey70") +
  ggplot2::geom_raster(data = dplyr::filter(cs1_pca_fpoints,
                                            pc == unique(pc)[1]),
                       ggplot2::aes(x = x, y = y, fill = Score)) +
  ggplot2::geom_polygon(data = cs1_pus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = NA, color = "black", size = 0.05) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white"),
    axis.title = ggplot2::element_blank(),
    legend.background = ggplot2::element_rect(color = "transparent",
                                              fill = "transparent"),
    legend.position = c(1.1, 1.07),
    legend.justification = c(1,1),
    legend.key.height = ggplot2::unit(1.5, "mm"),
    legend.key.width = ggplot2::unit(2, "mm")) +
  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
                           expand = FALSE) +
  ggplot2::scale_fill_continuous(breaks =
    pretty(dplyr::filter(cs1_pca_fpoints, pc == unique(pc)[1])$Score, n = 3,
          min.n = 2)) +
  ggplot2::facet_wrap(~ pc)
p2 <- ggplot2::ggplot() +
  ggplot2::geom_polygon(data = aus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill="grey85", color="grey70") +
  ggplot2::geom_raster(data = dplyr::filter(cs1_pca_fpoints,
                                            pc == unique(pc)[2]),
                       ggplot2::aes(x = x, y = y, fill = Score)) +
  ggplot2::geom_polygon(data = cs1_pus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = NA, color = "black", size = 0.05) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white"),
    axis.title = ggplot2::element_blank(),
    legend.background = ggplot2::element_rect(color="transparent",
                                              fill="transparent"),
    legend.position = c(1.1, 1.07),
    legend.justification = c(1,1),
    legend.key.height = ggplot2::unit(1.5, "mm"),
    legend.key.width = ggplot2::unit(2, "mm")) +
  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
                           expand = FALSE) +
  ggplot2::scale_fill_continuous(breaks =
    pretty(dplyr::filter(cs1_pca_fpoints, pc == unique(pc)[2])$Score, n = 2,
           min.n = 2)) +
  ggplot2::facet_wrap(~ pc)
# render plots
gridExtra::grid.arrange(p1, p2, ncol = 2)
```

```{r, echo = FALSE, fig.height = 3.8, fig.width = column_width, fig.cap = "Distribution of the species used in the first case study. Polygons denote planning units. Planning units occupied by a given species are shown in light blue."}
# prepare data
cs1_pus_occ_fpolygons <- plyr::ldply(
  seq_along(cs1_parameters$common_names),
  function(i) {
    curr_data <- cs1_pus_fpolygons
    curr_data$species <- cs1_parameters$common_names[i]
    curr_data$Occupancy <- "absent"
    curr_pus <- cs1_ru@data@pu.species.probabilities %>%
                dplyr::filter(species == i) %>%
                `[[`("pu")
    curr_data$Occupancy[which(curr_data$id %in% curr_pus)] <- "present"
    return(curr_data)
})

cs1_pus_occ_fpolygons$species <- factor(
    as.character(cs1_pus_occ_fpolygons$species),
    levels = cs1_parameters$common_names)
# make letter data.frame
cs1_pus_occ_letters_data <- data.frame(
  x = unname(cs1_pus@bbox[1,1] + (diff(cs1_pus@bbox[1, ]) * 0.08)),
  y = unname(cs1_pus@bbox[2,1] + (diff(cs1_pus@bbox[2, ]) * 0.95)),
  species=unique(cs1_pus_occ_fpolygons$species),
  letter=paste0("(", letters[seq_along(unique(cs1_pus_occ_fpolygons$species))],
                ")")
)
# make plots
ggplot2::ggplot() +
  ggplot2::geom_polygon(data = aus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = "grey85", color = "grey70", size = 0.05) +
  ggplot2::geom_polygon(data = cs1_pus_occ_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group,
                                     fill = Occupancy),
                        color = "black", size = 0.05) +
  ggplot2::geom_label(data = cs1_pus_occ_letters_data,
                      ggplot2::aes(label = letter, x = x, y = y),
                      color = "black", fill = scales::alpha("white", 0.8), size = 3) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white", size = 7),
    axis.title = ggplot2::element_blank(),
    legend.position = "none") +
  ggplot2::scale_fill_manual(name = "Occupancy",
                             values = c("absent"="#132B43",
                                        "present"="#56B1F7")) +
  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
                           expand = FALSE) +
  ggplot2::facet_wrap(~ species, ncol = 2)
```
